Experience shapes individual foraging specialization and success in a virtual predator-prey system

### 2024-09-06 manuscript/manuscript.Rmd
files up to date

The capacity of predators to match their tactic to the behaviour of their prey is expected to drive the outcome of predator-prey interactions. To do so, predators need to practice their tactics extensively to develop the skills and expertise to successfully hunt their prey, which may result in individual foraging specialization. Yet, there are very limited empirical assessments showing links between experience, foraging specialization, and hunting success at the individual level, due to the numerous challenges of monitoring direct interactions in the wild. In this study, we used a virtual predator-prey system (the online game *Dead by Daylight*) to investigate if and how hunting experience shapes individual foraging specialization and success in predators across repeated interactions with their prey. Our results show that predators optimized prey consumption as they gained experience. Prey behaviour was an important mediator of this relationship, resulting in differences among predators in the development of expertise. Throughout experience, 68% of predators became either increasingly specialized by always moving at a fast pace, or flexible by switching between slow and fast speeds. The remaining 32% of predators used a specialized tactic of fast movement throughout the study period. The predators' strategies (i.e. flexible or specialized foraging) were partially matched to the preys' speed, suggesting that changes in hunting behaviour were driven by repeated encounters with prey. Ultimately, flexible foragers were only slightly less successful than specialist foragers, such that both achieved similar success overall. Our observations suggest that experience drives among individual differences in foraging tactics, which may act as a stabilizing mechanism promoting behavioural diversification in predator-prey systems.

>>>>

The capacity of predators to match their tactic to their prey is expected to drive the outcome of predator-prey interactions. To do so, predators need to practice their tactics extensively to develop the skills to be successful hunters, which may result in individual foraging specialization. Yet, there are limited empirical assessments showing links between experience, foraging specialization, and hunting success at the individual level, due to the challenges of monitoring direct interactions in the wild. Here, we used a virtual predator-prey system (the game *Dead by Daylight*) to investigate how experience shapes individual foraging specialization and success in predators across repeated interactions with their prey. We show that predators optimized prey consumption as they gained experience. Prey speed was an important mediator of this relationship, resulting in differences among predators in the development of expertise. Throughout experience, 68% of predators became either increasingly specialized by always moving at a fast pace, or flexible by switching between slow and fast speeds. The predators’ strategies were partially matched to the preys’ speed, suggesting that changes in hunting behaviour were driven by repeated encounters with prey. Flexible and specialist foragers achieved similar success overall. Our findings suggest that experience may be mechanism promoting behavioural diversification in predator-prey systems.
